plyranges: a grammar for transforming genomics data

### 2018-04-13 paper.Rmd
moving things about

comparsion to Bioconductor API here?

@@

 (that
are named as nouns) with many arguments (which is required in the existing 
Bioconductor packages)
 This is possible because every 
function call is an endomorphism: when the input is GRanges object the output
will also be a GRanges object. One advantage of endomorphism is that it 
does not  require any additional learning of classes beyond GRanges and 
the _DataFrame_ classes. This strongly deviates from the design of 
the GRanges Bioconductor packages, where many methods return a new class
upon return. The Bioconductor design enables efficient computing as users
are exposed to low-level features of its API which _plyranges_ abstracts
away. Method chaining via the pipe operator can also be difficult to debug,
as there multiple points of failure.

@@

We have designed every function to perform a single
action on a GRanges object. The `plyranges` DSL implments the core verbs from the `dplyr` package
and implements context specific verbs for transforming GRanges objects (described in
table ). Each verb preserves the semantics of GRanges object and via
duck typing works with derivatives of the GRanges class. 
Both of these aspects reduce the cognitive load on a new user since most 
manipulations can be performed with a  vocabulary of several verbs, rather 
than having to memorise function names that are nouns. Moreover, this approach 
strongly constrasts the `GenomicRanges/IRanges` OO interface, which emphasises 
the use of setter and getter methods. In that interface, core components and 
metadata are updated via replacement methods (requiring knowledge of the
class components), while our interface requires only a single call to `mutate()`
to perform the modification.

@@

@@

>>>>

comparsion to Bioconductor API here? table of join ops?

@@

This is possible because every 
function call is an endomorphism: when the input is GRanges object the output
will also be a GRanges object. One advantage of endomorphism is that it 
does not  require any additional learning of classes beyond GRanges and 
the _DataFrame_ classes. This strongly deviates from the design of 
the GRanges Bioconductor packages, where many methods return a new class
upon return. The Bioconductor design enables efficient computing as users
are exposed to low-level features of its API which _plyranges_ abstracts
away.

@@

Using the design principles defined above every function in `plyranges`
performs a single action on GRanges objects. The `plyranges` DSL implments the 
core verbs from the `dplyr` package and implements a genomic relational algebra
for transforming GRanges objects (table x, y). Each verb preserves the 
semantics of GRanges object and works with derivatives of the GRanges class. 
Both of these aspects reduce the cognitive load on a new user since most 
manipulations can be performed with a vocabulary of several verbs, rather 
than having to memorise function names that are nouns.

This approach strongly constrasts the `GenomicRanges/IRanges` OO interface, 
which emphasises the use of setter and getter methods. In that interface, 
core components and metadata are updated via replacement methods 
(requiring knowledge of the class components), while our interface requires 
only a single call to `mutate()` to perform the modification.

@@

Method chaining via the pipe operator can also be difficult to debug,
as there multiple points of failure

@@

Here we provide examples on how to use the `plyranges` DSL to construct
genomic data workflows.
